Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-10-13 vesicle_code/docs/permeableVes/main.tex
starting contracting geometry runs
converted tikz to pdf

@@

\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/ellipseArea.tikz}
  \fi
%  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The time axis is scaled by the permeability rate $\beta$. Since
  the curves collapse to a single curve, the permeability rate $\beta$
  sets a relaxation time scale.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Shear Flow}
We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$. Depending on its reduced
area and viscosity contrast, two-dimensional vesicles either tank tread
or tumble~\cite{fin-lam-sei-gom2008,kra-win-sei-lip1996}. Since we are
considering semi-permeable vesicles, we assume there is no viscosity
contrast, and tumbling is not possible.
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/shearSS.tikz}
  \fi
  \caption{\label{fig:shearSS} The reduced area of a semi-permeable
  vesicle in a shear flow. The final reduced area is nearly independent
  of the initial reduced area.}
\end{figure}

@@

\begin{figure}[htp]
  \ifTikz
  \input{figures/parabolicOffCenterPhaseImpermeable.tikz}
  \fi
  \caption{\label{fig:parabolicOffCenterImpermeablePhaseDiagram} Phase
  diagram of an impermeable vesicle with several different flow rates
  $U$ and reduced areas $\nu$. As shown by Kaoui et
  al.~\cite{kao-bir-mis2009}, the final vesicle shape is either an
  axisymmetric parachute or bullet (pink region) or an asymmetric
  slipper (yellow region).}
\end{figure}

@@

\begin{figure}[htp]
  \ifTikz
  \input{figures/parabolicOffCenterPhaseSemipermeable.tikz}
  \fi
  \caption{\label{fig:parabolicOffCenterSemipermeablePhaseDiagram} Phase
  diagram of an semi-permeable vesicle, with $\beta = 10^{-3}$, with
  several different flow rates $U$ and reduced areas $\nu$. The
  steady state reduced area for each flow rate, which is reported along
  the top of the vesicles, is nearly independent of the initial reduced
  area. Low flow rates result in axisymmetric vesicles with large
  reduced area, and high flow rates result in asymmetric vesicles with
  low reduced area. Results at higher flow rates are not eporeted since
  the vesicle self-intersects.}
\end{figure}

\begin{itemize}
  \item If the flow rate is too small, the vesicle only migrates part
    way before it turns into a circle. Once it's a circle, it acts like
    a rigid body, so it no longer migrates towards the center line
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in Stenosis}
We consider the effect of stenosis on a semi-permeable vesicle. To allow
for large amounts of fluid flux through the membrane, we consider
geometry with a long constricted region. We start by considering a
vesicle initialized with reduced area $\nu = 0.65$, and we compare the
vesicle shape when $\beta = 0$ (impermeable) and $\beta = 10^{-3}$
(permeable). Figure~\ref{fig:stenosisShape} shows the vesicles' shape at
several locations along the constriction. The colors along the left
channel represent the tension and those along the right represent the
flux. In Figure~\ref{fig:stenosisTensionAndFlux}, the tension at
location $x=20$ and the flux at location $x=30$ are plotted with respect
to arclength. The value of $\theta=0$ corresponds to the nose of the
vesicle.
\begin{figure}[htp]
  \ifTikz
%  \scalebox{0.5}{\input{figures/stenosisShapeTension01.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux02.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension03.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux04.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension05.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux06.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension07.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux08.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension09.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux10.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension11.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux12.tikz}}
  \scalebox{0.5}{\input{figures/stenosisShapeTension02.tikz}}
  \scalebox{0.5}{\input{figures/stenosisShapeFlux04.tikz}}
  \scalebox{0.5}{\input{figures/stenosisShapeTension06.tikz}}
  \scalebox{0.5}{\input{figures/stenosisShapeFlux08.tikz}}
  \scalebox{0.5}{\input{figures/stenosisShapeTension10.tikz}}
  \scalebox{0.5}{\input{figures/stenosisShapeFlux12.tikz}}
  \fi
  \caption{\label{fig:stenosisShape} An impermeable (dashed) and
  semi-permeable (solid) vesicle in a stenosed geometry (black).
  The vesicles' initial configuration are identical with reduced area
  $\nu = 0.65$. The color denotes the vesicles' tension (left) and flux
  (right). The semi-permeable vesicle deflates throughout the
  constriction (see Figure~\ref{fig:stenosisArea}) and develops higher
  curvature regions than its impermeable counterpart.}
\end{figure}

\begin{figure}[htp]
  \centering
  \ifTikz
  \scalebox{0.5}{\input{figures/stenosisTension12.tikz}}
  \scalebox{0.5}{\input{figures/stenosisFlux12.tikz}}
  \scalebox{0.5}{\input{figures/stenosisXvelocity12.tikz}}
  \scalebox{0.5}{\input{figures/stenosisYvelocity12.tikz}}
  \fi
  \caption{\label{fig:stenosisTensionAndFlux} The tension (top left), 
  flux (top right), $x$-velocity (bottom left), and $y$-velocity (bottom
  right) of the semi-permeable (red) and impermeable (dashed blue)
  vesicles in Figure~\ref{fig:stenosisShape} at location $x=50$. The
  point $\theta=0$ corresponds to the right-most point of the vesicle.}
\end{figure}

\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/stenosisArea.tikz}
  \fi
%  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:stenosisArea} The area of a stenosed
  semi-permeable vesicle, with permeability rate $\beta=10^{-3}$. The
  $x$-axis is the $x$-coordinate of the center of mass of the vesicle.
  The vesicle loses fluid at a constant rate throughout the constriction
  that is between $\pm 60$. The semi-permeable vesicle loses 12.8\% of
  its fluid.}
\end{figure}

\begin{figure}[htp]
%  \ifTikz
  \input{figures/stenosisBL.tikz}
%  \fi
  \caption{\label{fig:stenosisBL} A stenosed semi-impermeable (red) and
  imperemable (blue) vesicle. The semi-permeable vesicle has 
  reduced area 0.17, and both vesicles were initialized with reduce area
  $0.20$. The bottom plot shows the velocity relative to the vesicle
  inside the lubrication layer.  The vesicle velocity in the square is
  nearly constant (see Figure~\ref{fig:stenosisRA020Velocity}.}
\end{figure}

\begin{figure}[htp]
  \centering
  \ifTikz
  \scalebox{0.5}{\input{figures/stenosisXvelocityRA020.tikz}}
  \scalebox{0.5}{\input{figures/stenosisYvelocityRA020.tikz}}
  \fi
  \caption{\label{fig:stenosisRA020Velocity} The $x$ (left) and $y$
  (right) velocities of the vesicle in Figure~\ref{fig:stenosisBL}. The
  red lines correspond to the semi-permeable vesicle and the blue dashed
  lines correspond to the impermeable vesicle.}
\end{figure}

@@

\begin{figure}[htp]
  \ifTikz
  \input{figures/stenosisGeom.tikz}
  \fi
%  \includegraphics[width=0.8\linewidth]{figures/parabolicPhaseDiagram.pdf}
  \caption{\label{fig:stenosisGeom} A single semi-permeable vesicle
  passing through a constriction. The geometry and parameters are chosen
  to replicate the experiments of Abkarian et
  al.~\cite{abk-fai-sto2006}.}
\end{figure}

\begin{figure}[htp]
  \ifTikz
  \input{figures/stenosisPressureFluctuation.tikz}
  \fi
  \caption{\label{fig:stenosisPressure} The excess pressure across the
  constriction in Figure~\ref{fig:stenosisGeom}. The black dashed lines
  denote the start and end of the constriction. The results are
  qualitative agreement with~\cite{abk-fai-sto2006}.}
\end{figure}

>>>>

\paragraph{Literature to maybe cite}
\cite{ali-ovr-ran2020} show how beading can occur on nanotubes. This
happens in regions where proteins congregate and this can lead to
semi-permeability?

@@

\begin{figure}[htp]
  \centering
%  \ifTikz
%  \input{figures/ellipseArea.tikz}
%  \fi
  \includegraphics[width=0.9\linewidth]{figures/ellipseArea}
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The time axis is scaled by the permeability rate $\beta$. Since
  the curves collapse to a single curve, the permeability rate $\beta$
  sets a relaxation time scale.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Shear Flow}
We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$. Depending on its reduced
area and viscosity contrast, two-dimensional vesicles either tank tread
or tumble~\cite{fin-lam-sei-gom2008,kra-win-sei-lip1996}. Since we are
considering semi-permeable vesicles, we assume there is no viscosity
contrast, and tumbling is not possible.
\begin{figure}[htp]
  \centering
%  \ifTikz
%  \input{figures/shearSS.tikz}
%  \fi
  \includegraphics[width=0.9\linewidth]{figures/shearSS.pdf}
  \caption{\label{fig:shearSS} The reduced area of a semi-permeable
  vesicle in a shear flow. The final reduced area is nearly independent
  of the initial reduced area.}
\end{figure}

@@

\begin{figure}[htp]
%  \ifTikz
%  \input{figures/parabolicOffCenterPhaseImpermeable.tikz}
%  \fi
  \includegraphics[width=0.9\linewidth]{figures/parabolicOffCenterPhaseImpermeable.pdf}
  \caption{\label{fig:parabolicOffCenterImpermeablePhaseDiagram} Phase
  diagram of an impermeable vesicle with several different flow rates
  $U$ and reduced areas $\nu$. As shown by Kaoui et
  al.~\cite{kao-bir-mis2009}, the final vesicle shape is either an
  axisymmetric parachute or bullet (pink region) or an asymmetric
  slipper (yellow region).}
\end{figure}

@@

\begin{figure}[htp]
%  \ifTikz
%  \input{figures/parabolicOffCenterPhaseSemipermeable.tikz}
%  \fi
  \includegraphics[width=0.9\linewidth]{figures/parabolicOffCenterPhaseSemipermeable.pdf}
  \caption{\label{fig:parabolicOffCenterSemipermeablePhaseDiagram} Phase
  diagram of a semi-permeable vesicle, with $\beta = 10^{-3}$, with
  several different flow rates $U$ and reduced areas $\nu$. The
  steady state reduced area for each flow rate, which is reported along
  the top of the vesicles, is nearly independent of the initial reduced
  area. Low flow rates result in axisymmetric vesicles with large
  reduced area, and high flow rates result in asymmetric vesicles with
  low reduced area. Results at higher flow rates are not eporeted since
  the vesicle self-intersects.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in Stenosis}
We consider the effect of stenosis on a semi-permeable vesicle. To allow
for large amounts of fluid flux through the membrane, we consider
geometry with a long constricted region. We start by considering a
vesicle initialized with reduced area $\nu = 0.65$, and we compare the
vesicle shape when $\beta = 0$ (impermeable) and $\beta = 10^{-3}$
(permeable). Figure~\ref{fig:stenosisShape} shows the vesicles' shape at
several locations along the constriction. The colors along the left
channel represent the tension and those along the right represent the
flux. In Figure~\ref{fig:stenosisTensionAndFlux}, the tension at
location $x=20$ and the flux at location $x=30$ are plotted with respect
to arclength. The value of $\theta=0$ corresponds to the nose of the
vesicle.
\begin{figure}[htp]
%  \ifTikz
%  \scalebox{0.5}{\input{figures/stenosisShapeTension01.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux02.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension03.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux04.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension05.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux06.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension07.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux08.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension09.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux10.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension11.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux12.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension02.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux04.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension06.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux08.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeTension10.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisShapeFlux12.tikz}}
%  \fi
  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeTension02.pdf}
  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeFlux04.pdf}
  \\
  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeTension06.pdf}
  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeFlux08.pdf}
  \\
  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeTension10.pdf}
  \includegraphics[width=0.45\linewidth]{figures/stenosisShapeFlux12.pdf}
  \caption{\label{fig:stenosisShape} An impermeable (dashed) and
  semi-permeable (solid) vesicle in a stenosed geometry (black).
  The vesicles' initial configuration are identical with reduced area
  $\nu = 0.65$. The color denotes the vesicles' tension (left) and flux
  (right). The semi-permeable vesicle deflates throughout the
  constriction (see Figure~\ref{fig:stenosisArea}) and develops higher
  curvature regions than its impermeable counterpart.}
\end{figure}

\begin{figure}[htp]
  \centering
%  \ifTikz
%  \scalebox{0.5}{\input{figures/stenosisTension12.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisFlux12.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisXvelocity12.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisYvelocity12.tikz}}
%  \fi
  \includegraphics[width=0.45\linewidth]{figures/stenosisTension12.pdf}
  \includegraphics[width=0.45\linewidth]{figures/stenosisFlux12.pdf} \\
  \includegraphics[width=0.45\linewidth]{figures/stenosisXvelocity12.pdf}
  \includegraphics[width=0.45\linewidth]{figures/stenosisYvelocity12.pdf}
  \caption{\label{fig:stenosisTensionAndFlux} The tension (top left), 
  flux (top right), $x$-velocity (bottom left), and $y$-velocity (bottom
  right) of the semi-permeable (red) and impermeable (dashed blue)
  vesicles in Figure~\ref{fig:stenosisShape} at location $x=50$. The
  point $\theta=0$ corresponds to the right-most point of the vesicle.}
\end{figure}

\begin{figure}[htp]
  \centering
%  \ifTikz
%  \input{figures/stenosisArea.tikz}
%  \fi
  \includegraphics[width=0.8\linewidth]{figures/stenosisArea}
  \caption{\label{fig:stenosisArea} The area of a stenosed
  semi-permeable vesicle, with permeability rate $\beta=10^{-3}$. The
  $x$-axis is the $x$-coordinate of the center of mass of the vesicle.
  The vesicle loses fluid at a constant rate throughout the constriction
  that is between $\pm 60$. The semi-permeable vesicle loses 12.8\% of
  its fluid.}
\end{figure}

\begin{figure}[htp]
%  \ifTikz
%  \input{figures/stenosisBL.tikz}
%  \fi
  \includegraphics[width=\linewidth]{figures/stenosisBL.pdf}
  \caption{\label{fig:stenosisBL} A stenosed semi-impermeable (red) and
  imperemable (blue) vesicle. The semi-permeable vesicle has 
  reduced area 0.17, and both vesicles were initialized with reduce area
  $0.20$. The bottom plot shows the velocity relative to the vesicle
  inside the lubrication layer.  The vesicle velocity in the square is
  nearly constant (see Figure~\ref{fig:stenosisRA020Velocity}.}
\end{figure}

\begin{figure}[htp]
  \centering
%  \ifTikz
%  \scalebox{0.5}{\input{figures/stenosisXvelocityRA020.tikz}}
%  \scalebox{0.5}{\input{figures/stenosisYvelocityRA020.tikz}}
%  \fi
  \includegraphics[width=0.45\linewidth]{figures/stenosisXvelocityRA020.pdf}
  \includegraphics[width=0.45\linewidth]{figures/stenosisYvelocityRA020.pdf}
  \caption{\label{fig:stenosisRA020Velocity} The $x$ (left) and $y$
  (right) velocities of the vesicle in Figure~\ref{fig:stenosisBL}. The
  red lines correspond to the semi-permeable vesicle and the blue dashed
  lines correspond to the impermeable vesicle.}
\end{figure}

@@

\begin{figure}[htp]
%  \ifTikz
%  \input{figures/stenosisGeom.tikz}
%  \fi
  \includegraphics[width=\linewidth]{figures/stenosisGeom.pdf}
  \caption{\label{fig:stenosisGeom} A single semi-permeable vesicle
  passing through a constriction. The geometry and parameters are chosen
  to replicate the experiments of Abkarian et
  al.~\cite{abk-fai-sto2006}.}
\end{figure}

\begin{figure}[htp]
%  \ifTikz
%  \input{figures/stenosisPressureFluctuation.tikz}
%  \fi
  \includegraphics[width=\linewidth]{figures/stenosisPressureFluctuation.pdf}
  \caption{\label{fig:stenosisPressure} The excess pressure across the
  constriction in Figure~\ref{fig:stenosisGeom}. The black dashed lines
  denote the start and end of the constriction. The results are
  qualitative agreement with~\cite{abk-fai-sto2006}.}
\end{figure}
